---
title: "AI for Statistical Learning — Week 1"
subtitle: "Week 1 Report: Optimizing Linear Regression using Distribution Theory"
author: "Group 2: Carlos, Landon, and Elias."
date: 2026-06-06

format:
    html:
        papersize: letter
        fontsize: 11pt
        geometry: margin=1in
        number-sections: true
        toc: false
    pdf:
        pdf-engine: xelatex
        embed-resources: true
        papersize: letter
        fontsize: 11pt
        geometry: margin=1in
        number-sections: true
        toc: false
        include-in-header:
            text: |
                \usepackage{fvextra}
                \usepackage{xurl}
                \fvset{breaklines=true, breakanywhere=true}
                \setlength{\emergencystretch}{3em}
                \sloppy
---

:::{.callout-note collapse="true"}

## Introduction

Previously, we wrote our linear regression model as $y_i = \beta_0 + \beta_1 x_i + \epsilon_i$,  

where $\epsilon_i \sim N(0, \sigma^2)$. 

Turn out, using the properties of the normal distribution, an equivalent specification is:
$y_i \sim N(\beta_0 + \beta_1 x_i, \sigma^2)$. 

So instead of using $\sum \epsilon_i^2$ as the loss function, we can also use the negative log likelihood of the normal distribution as the loss function.


Consider the following data generation process:



```{r}
set.seed(0)
n <- 1000
x <- rnorm(n = n, mean = 0, sd = 2)
b0 <- -2; b1 <- 7.7
y <- rnorm(n, b0 + b1 * x, sd = 6)
```


The likelihood of the normal distribution for $y_i$, ..., $y_n$ (where n = 1000) is: 

<div style="text-align: center;">
$L(\beta | y) = \frac{1}{(6 \sqrt{2\pi})^n}\exp \biggl\{-\frac{1}{2 \cdot 6^2} \sum_{i=1}^n (y_i - \beta_0 - \beta_1x_i)^2\biggr\}$

</div>

where $\beta = (\beta_0, \beta_1)$ 


Estimate $\beta$ using the following steps: 

1. Write down an expression of the negative log-likelihood (nll) - $-l(\beta | y) = - log L(\beta | y)$ 

2. Use the `optim` function to minimize the nll and estimate $\beta$. 

3. Confirm your answer by using the `dnorm` function instead of manually calculating the nll. 

4. [BONUS] In the previous parts, we fixed $\sigma = 6$. Can you also try estimating it from the data as well? You will need to estimate $\theta = (\beta_0, \beta_1, \sigma)$ where the first components are unbounded but the last one is strictly positive.

:::

:::{.callout-note collapse="true"}

## Answer 1

- The likelihood of the normal Distribution: 

<div style="text-align: center;">
$L(\beta | y) = \frac{1}{(\sigma \sqrt{2\pi})^n}\exp \biggl\{-\frac{1}{2 \cdot \sigma^2} \sum_{i=1}^n (y_i - \beta_0 - \beta_1x_i)^2\biggr\}$

</div>

- The expression of the negative log-likelihood (nll): 
<div style="text-align: center;">
$nll = - l(\beta | y) = - \log L(\beta | y)$ ${\large \Rightarrow } - l(\beta | y) = - \log \biggl( \frac{1}{(\sigma \sqrt{2\pi})^n} \exp \biggl\{-\frac{1}{2 \cdot \sigma^2} \sum_{i=1}^n (y_i - \beta_0 - \beta_1 x_i)^2\biggr\} \biggr)$
</div>

- Using some properties of logarithms, we obtain the following simplified expression for the nll:
<div style="text-align: center;">
$nll = - l(\beta\mid y) = + n \log\bigl(\sigma \sqrt{2\pi}\bigr) + \frac{1}{2\sigma^2} \sum_{i=1}^n (y_i - \beta_0 - \beta_1 x_i)^2$. 
</div>

- In our case, since $\sigma = 6$, and $n = 1000$, the expression becomes:
<div style="text-align: center;">
$nll = 1000 \log\bigl(6 \sqrt{2\pi}\bigr) + \frac{1}{72} \sum_{i=1}^{1000} (y_i - \beta_0 - \beta_1 x_i)^2$.
</div>

:::


:::{.callout-note collapse="true"}

## Answer 2

- Write the nll function to be minimized by `optim` and then minimize it:
```{r}
nll_loss <- function(theta){
  b0 = theta[1]
  b1 = theta[2]
  nll = n * log(6*sqrt(2*pi)) + sum((y - b0 - b1*x)^2)/72
  return(nll)
}
```

- Minimizing the nll function:
```{r}
optim(c(0, 0), nll_loss)
```

:::

:::{.callout-note collapse="true"}

## Answer 3
- Confirming the answer using the `dnorm` function instead of manually calculating the nll:
```{r}
ll <- function(theta){
  b0 = theta[1]
  b1 = theta[2]
  # Sum of the logs of the density at each point in the y vector
  llik = sum(dnorm(y, mean = b0 + b1 * x, log = T))
  # Negative log-likelihood to be minimized
  return(-llik)
}
```
- Minimizing the nll function:
```{r}
optim(c(0, 0), ll)
```

:::

:::{.callout-note collapse="true"}

## Bonus Question
```{r}
nll_Sig <- function(theta){
  b_0 = theta[1]
  b_1 = theta[2]
  sigma = theta[3]
  nll = n * log(sigma*sqrt(2*pi)) + sum((y - b_0 - b_1*x)^2)/(2*(sigma)^2)
  return(nll)
}
```

- Minimizing the nll function:

```{r}
optim(c(0, 0, 1),
      nll_Sig,
      method = "L-BFGS-B",
      lower = c(-Inf, -Inf, 1e-10),
      upper = c(Inf, Inf, Inf))
```
:::
