---
title: "University of Arkansas STAT6393V - Week 5"
subtitle: "Week 5 - Final Project: Estimation and Uncertainty Quantification of rainfall data."
author: "Group 2: Carlos, Landon, and Elias."
date: 2026-06-18

format:
    html:
        papersize: letter
        fontsize: 12pt
        geometry: margin=1in
        number-sections: true
        toc: false
    pdf:
        pdf-engine: xelatex
        embed-resources: true
        papersize: letter
        fontsize: 11pt
        geometry: margin=1in
        number-sections: true
        toc: false
        include-in-header:
            text: |
                \usepackage{fvextra}
                \usepackage{xurl}
                \fvset{breaklines=true, breakanywhere=true}
                \setlength{\emergencystretch}{3em}
                \sloppy
---


:::{.callout-note collapse="true"}

## Introduction

In general, datsets are made of variables of different data types. The value of the data for these variables can take values that are discrete, continuous, or both (semi-continuous). When a variable includes both discrete and continuous values (components), it is referred to as semi-continuous data type. The choice of a model to represent the distribution of a variable in a dataset is driven by the type of data contained within that variable. For example, if the variable's data type is continuous, we can use a normal distribution model to represent its distribution. If the variable's data type is discrete, we can use a Poisson distribution model. However, if the data contained within the variable include both discrete and continuous components, we can use a semi-continuous mixture model to represent its distribution.

In this project, we use the weather dataset which includes a varible for maximum temperature (tmax), and a variable for precipitation (pr). The maximum temperature variable is continuous, as it can take on any value within a certain range. The precipitation variable on the other hand provides the amount of rainfall in a given location and time, reflecting the possibility of both no precipitation (discrete) and positive precipitation (continuous) values, making the data type of this variable semi-continuous.


The main objectives of this project are: 

1. Model the conditional distribution of precipitation given maximum temperature (pr | tmax).
2. Answer the following scientific questions: 

    - Does the probability of rain depend on the temperature?
    - Does the amount of rain depend on the temperature?


:::


:::{.callout-note collapse="true" fontsize="large"}

## Part 1: Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cito)
library(torch)
library(lubridate)
library(SPQR)
#source('BNN_layers.R') #By Putting this here, we don't need to define functions again
```

In this first step of the project, we load all necesary libraries, load and explore the weather dataset, prepare it for analysis according to the directions assigned to our group (Group 2), and then define a custom loss function for a semi-continuous gamma mixture model.

**1. Data Exploration and Preparation** 

We first load the dataset and explore its structure and dimensions. 

```{r load dataset and explore it, include=FALSE}
weather <- readRDS("weather.RDS")
dim(weather)
str(weather)
maxt <- max(weather$time)
mint <- min(weather$time)

maxt
mint

maxtmax <- max(weather$tmax)
mintmax <- min(weather$tmax)
maxtmax
mintmax

maxpr <- max(weather$pr)
minpr <- min(weather$pr)
maxpr
minpr
```


The weather dataset contains daily weather data from January 1, 1951 to December 31, 2014. Its structure has 7 variables (columns) and 584,000 observations (rows). The variables are: time, latitude, longitude, maximum temperature (tmax), precipitation (pr), month, and location. The dataset is in a long format, with each row representing a single observation for a specific location and time. 

The month assigned to our group (Group 2) was June. We first subset the dataset to only include the observations for June and then convert all temperatures from Kelvin (K) to degrees Celsius (°C). We also set all precipitation values less than 0.01 to 0.



```{r Select a subset of the dataset, include=FALSE}
data <- weather[weather$month == 6, ]
dim(data)
head(data)
```


```{r Convert temperature values from Kelvin (K) to degrees Celsius (°C), include=FALSE}
data$tmax <- data$tmax - 273.15
str(data)
```


```{r Set all precipitation values less than 0.01 to 0, include=FALSE}
data$pr[data$pr < 0.01] <- 0
str(data)
```

The resulting subset dataset contains 48,000 observations and 7 variables.


```{r training and testing sets split, include=FALSE}
train_rows <- which(year(data$time) > 1960 & year(data$time) < 2001)
train_set <- data[train_rows, ]
test_set <- data[-train_rows, ]

dim(train_set)
head(train_set)

dim(test_set)
head(test_set)
```

We split the subset dataset into training and testing sets, we will use the data from 1961 to 2000 as the training set and the data from 1951 to 1960 and 2001 to 2014 as the testing set. The training set contains 30,000 observations, while the testing set contains 18,000 observations. 

The true split is as follows: 

    - Training set: 30,000 / 48,000 = 0.625 = 62.5%
    - Testing set: 18,000 / 48,000 = 0.375 = 37.5% 

This split is approximatly 60% for training and 40% for testing, allowing us to train our model on a substantial amount of data while reserving a significant portion for evaluating its performance on unseen data. 


**2. Custom Loss Function for semi-continuous gamma mixture:**

Now that our dataset is ready for analysis, we define a custom loss function for a semi-continuous gamma mixture model. This loss function will be used to fit the model to the training data and evaluate its performance on the testing data.

- The Loss Function for semi-continuous gamma mixture model:


Taking into account the mixture model:

$$Y\ |\ X, \theta \sim \ \pi_0 \ + \ (1 - \pi_0)Gamma(\alpha, \beta)$$

Here, $\theta = (\pi_0, \alpha, \beta)$, hence the parameters to be estimated were therefore $\pi_0$, $\alpha$ and $\beta$. Here, $\pi_0$ represents the probability that it will not rain, whereas the $Gamma$ distribution portion of the formula models precipitation in terms of temperature if it rains. $\alpha$ and $\beta$ are parameters of the $Gamma$ distribution, relevant for estimating the mean precipitation.

Importantly, the $Gamma$ portion of the formula is only relevant if it rains, hence the log-likelihood function must correctly account for that. The log-likelihood function is thus:

$$\begin{equation*}
l(\theta|Y) = \begin{cases} 
\sum\log\pi_0 & \text{if } y = 0 \\ 
\sum\log(1-\pi_0)\ + \ \sum\log(Gamma(\alpha, \beta) & \text{if } y > 0 
\end{cases}
\end{equation*}$$

**WE NEED TO FIX THE SUMS LIMITS IN THE LOG-LIKELIHOOD FUNCTION ABOVE.**



```{r, include=FALSE}
GSCMM_nll <- function(pred,true,...){
  yobs <- true[,1, drop = FALSE] # Extract Observed precipitation values
  rainy <- ifelse(yobs > 0, 1, 0) # Create a binary variable indicating whether it rained or not
  
  pi0 <- torch_sigmoid(pred[, 1, drop = FALSE]) # Extract the predicted probability of no rain
  alpha <- torch_exp(pred[, 2, drop = FALSE]) + 1e-6 # Extract the predicted shape parameter for the gamma distribution
  beta <- torch_exp(pred[, 3, drop = FALSE]) + 1e-6 # Extract the predicted rate parameter for the gamma distribution
  
  alpha <- alpha[rainy == 1] # Keep only the shape parameters for the rainy days
  beta <- beta[rainy == 1] # Keep only the scale parameters for the rainy days

# Log-likelihood split into two parts
# The first for non-rainy days, the second for rainy days

  ll <- sum(torch_log(pi0[rainy == 0])) +
        sum(torch_log(1 - pi0[rainy == 1]) + torch::distr_gamma(alpha,beta)$log_prob(yobs[rainy == 1])) # Calculate the log-likelihood for the semi-continuous gamma mixture model
  # Return the negative log-likelihood, which is the loss that will be minimized during model fitting
  return(-ll) # Return the negative log-likelihood.
}
```

The script for this custom loss function takes in the predicted values and the true values, and calculates the negative log-likelihood for the semi-continuous gamma mixture model. The function first extracts the observed precipitation values and creates a binary variable indicating whether it rained or not. It then extracts the predicted probability of no rain, as well as the predicted shape and rate parameters for the gamma distribution. Finally, it calculates the log-likelihood for the semi-continuous gamma mixture model and returns its negative value, which is equal to (-ll) the loss that will be minimized during model fitting.
:::

:::{.callout-note collapse="true"}

## Part 2: Model Fitting

**1. Fitting a model for pr | tmax:**

 In this second part of the project, we fit a model for precipitation as a function of maximum temperature at a location of our choice which we decided to be location 1. This subsetting operation resulted in a training dataframe with 1,200 observations and a testing dataframe with 720 observations. This number of observations may be small, but it is sufficient to fit a model for our analysis. We will use the training dataframe to fit the model and the testing dataframe to evaluate its performance. The results may not be as accurate as if we had used the entire dataset, but they will still provide valuable insights into the relationship between maximum temperature and precipitation.
 

```{r location 1 subset, include=FALSE}
train_loc1 <- train_set[train_set$loc == 1, ]
test_loc1 <- test_set[test_set$loc == 1, ]
dim(train_loc1)
dim(test_loc1)


train_df <- data.frame(x = train_loc1$tmax, y = train_loc1$pr)
test_df <- data.frame(x = test_loc1$tmax, y = test_loc1$pr)

dim(train_df)
dim(test_df)
```

We decided to fit multiple different neural network models for the semi-continuous gamma mixture model. We vary the number of hidden layers, the number of neurons in each layer, the epoches, and the batchsize, to find the best model for our data. It was a very computationally expensive experiment but we managed to complete it. 


```{r}
# | fig-high: 12
# | fig-width: 6

set.seed(007)
par(mfrow = c(1, 1))

model1 <- cito::dnn(
  cbind(y,y,y) ~ x, data = train_df, loss = GSCMM_nll,
  hidden = c(50, 50), activation = "relu",
  epochs = 100, lr = 0.001, batchsize = 25,
  optimizer = "adam", validation = 0.2,
  verbose = F, plot = FALSE
  )
cito:::visualize.training(
    losses = model1$losses,
    epoch = nrow(model1$losses),
    main = "Training Gamma Mixture Model 1",
    new = TRUE
)

model2 <- cito::dnn(
  cbind(y,y,y) ~ x, data = train_df, loss = GSCMM_nll,
  hidden = c(50, 50, 50), activation = "relu",
  epochs = 100, lr = 0.001, batchsize = 25,
  optimizer = "adam", validation = 0.2,
  verbose = F, plot = FALSE
  )
cito:::visualize.training(
    losses = model2$losses,
    epoch = nrow(model2$losses),
    main = "Training Gamma Mixture Model 2",
    new = TRUE
)


model3 <- cito::dnn(
  cbind(y,y,y) ~ x, data = train_df, loss = GSCMM_nll,
  hidden = c(100, 100), activation = "relu",
  epochs = 300, lr = 0.001, batchsize = 10,
  optimizer = "adam", validation = 0.2,
  verbose = F, plot = FALSE
  )
cito:::visualize.training(
    losses = model3$losses,
    epoch = nrow(model3$losses),
    main = "Training Gamma Mixture Model 3",
    new = TRUE
)

model4 <- cito::dnn(
  cbind(y,y,y) ~ x, data = train_df, loss = GSCMM_nll,
  hidden = c(100, 100, 100), activation = "relu",
  epochs = 100, lr = 0.001, batchsize = 12,
  optimizer = "adam", validation = 0.2,
  verbose = F, plot = FALSE
  )
cito:::visualize.training(
    losses = model4$losses,
    epoch = nrow(model4$losses),
    main = "Training Gamma Mixture Model 4",
    new = TRUE
)

# Save these plots as images for later use in the report
dev.copy(png, "model1_training_plot.png")
dev.copy(png, "model2_training_plot.png")
dev.copy(png, "model3_training_plot.png")
dev.copy(png, "model4_training_plot.png")
dev.off() # Close the graphics device to save the plots

```

Using the log-likelihood, we evaluated all four fitted models on the testing set. Because each model is evaluated on the same testing set, the best model is the one with the largest test-set log-likelihood, which is equivalent to the smallest test-set negative log-likelihood.

```{r}
models <- list(
  model1 = list(fit = model1, hidden_layers = 2, neurons = 50, epochs = 100, batchsize = 25),
  model2 = list(fit = model2, hidden_layers = 3, neurons = 50, epochs = 100, batchsize = 25),
  model3 = list(fit = model3, hidden_layers = 2, neurons = 100, epochs = 300, batchsize = 10),
  model4 = list(fit = model4, hidden_layers = 3, neurons = 100, epochs = 100, batchsize = 12)
)

calc_test_log_likelihood <- function(fit, data) {
  raw_pred <- predict(fit, newdata = data, type = "link")
  y <- data$y

  pi0 <- plogis(raw_pred[, 1])
  pi0 <- pmin(pmax(pi0, .Machine$double.eps), 1 - .Machine$double.eps)

  alpha <- exp(pmax(pmin(raw_pred[, 2], 50), -50))
  beta <- exp(pmax(pmin(raw_pred[, 3], 50), -50))

  log_lik <- numeric(length(y))
  dry <- y == 0
  rainy <- y > 0

  log_lik[dry] <- log(pi0[dry])
  log_lik[rainy] <- log1p(-pi0[rainy]) +
    dgamma(y[rainy], shape = alpha[rainy], rate = beta[rainy], log = TRUE)

  sum(log_lik)
}

model_selection <- do.call(rbind, lapply(names(models), function(m) {
  fit <- models[[m]]$fit
  losses <- fit$losses

  data.frame(
    model = m,
    hidden_layers = models[[m]]$hidden_layers,
    neurons_per_layer = models[[m]]$neurons,
    epochs = models[[m]]$epochs,
    batchsize = models[[m]]$batchsize,
    min_training_nll = min(losses$train_l, na.rm = TRUE),
    min_validation_nll = min(losses$valid_l, na.rm = TRUE),
    best_epoch = losses$epoch[which.min(losses$valid_l)],
    test_log_likelihood = calc_test_log_likelihood(fit, test_df)
  )
}))

model_selection$test_nll <- -model_selection$test_log_likelihood
model_selection$mean_test_nll <- model_selection$test_nll / nrow(test_df)
model_selection <- model_selection[order(model_selection$test_nll), ]
rownames(model_selection) <- NULL

best_model <- model_selection[1, ]
best_model_name <- best_model$model
model <- models[[best_model_name]]$fit

selection_sentence <- sprintf(
  paste0(
    "The best model is %s, with %s hidden layers, %s neurons per layer, ",
    "%s epochs, and a batch size of %s. It was selected because it had ",
    "the smallest test-set negative log-likelihood (%.2f), which means ",
    "it assigned the highest probability to the observed precipitation ",
    "values in the testing set."
  ),
  best_model$model,
  best_model$hidden_layers,
  best_model$neurons_per_layer,
  best_model$epochs,
  best_model$batchsize,
  best_model$test_nll
)

knitr::kable(
  model_selection,
  digits = 3,
  caption = "Model comparison using test-set log-likelihood. Models are ranked from best to worst by test negative log-likelihood."
)
```

`r selection_sentence` This fitted model is used for the validation plots and scientific questions below.




- Interpretations:

When the red line (validation) is below the training line (blue), it indicates that the model is overfitting the training data. This means that the model is performing well on the training data but not generalizing well to unseen data, which can lead to poor performance on the testing set. 

When the red line (validation) is above the training line (blue), it indicates that the model is underfitting the training data. This means that the model is not capturing the underlying patterns in the data and is not performing well on either the training or testing sets.


when fitting the models, the validation argument is set to 0.2 which means that 20% of the training data is used for validation during training. This allows us to monitor the model's performance on unseen data and helps to prevent overfitting. The batchsize argument is set to 50, which means that the model is trained on mini-batches of 100 observations at a time. This can help to speed up training and improve convergence.

epochs argument is set to 50, which means that the model is trained for 50 iterations over the entire training dataset. This can help to ensure that the model has enough time to learn the underlying patterns in the data.

to balance the batchsize and epochs arguments, we can adjust the batchsize to a smaller value (e.g., 25) and increase the number of epochs (e.g., 100). This will allow the model to see more data during training while still allowing for enough iterations to learn the underlying patterns in the data.


cbind(y, y, y) ~ x is used to specify the model formula for the neural network. It indicates that we want to predict three outputs (y, y, y) based on a single input variable (x). The cbind function is used to combine the three output variables into a matrix, which is required by the cito::dnn function. This allows us to fit a multi-output neural network model that can predict multiple outputs simultaneously.

the three outputs (y, y, y) represent the three parameters of the semi-continuous gamma mixture model: the probability of no rain (pi0), the shape parameter of the gamma distribution (alpha), and the scale parameter of the gamma distribution (beta). By predicting all three parameters simultaneously, we can capture the complex relationship between maximum temperature and precipitation in a single model.

to use the test data to evaluate the performance of the fitted models, we can use the predict function to generate predictions for the test set. We can then calculate the root mean squared error (RMSE) for each model to compare their performance. The model with the lowest RMSE on the test set can be considered the best model for predicting precipitation based on maximum temperature. 


include = FALSE is used to prevent the code chunk from being included in the final report. This is useful when we want to run code for testing or debugging purposes but do not want to include it in the final report. In this case, we are using include = FALSE for the model fitting and prediction code because we do not want to include it in the final report at all, as it is not relevant to the analysis.


 is done to reduce the complexity of the model and to focus on a specific location for analysis. We will then use the fitted model to predict precipitation values for a range of maximum temperature values and evaluate the model's performance using the testing set.



the difference between echo = false and include = false is that echo = false suppresses the output of the code chunk in the final report, while include = false completely excludes the code chunk from the final report. This means that the code will not be executed and will not produce any output or results. In this case, we are using include = false for the model fitting and prediction code because we do not want to include it in the final report at all, as it is not relevant to the analysis.

:::

:::{.callout-note collapse="true"}

## Part 3: Model Validation

To evaluate the model's ability to predict whether it will rain or not (the probability of not raining $\pi_0$), the observations in the testing set with precipitation superior to 0 were coded as 1, while the others were left as 0. $\pi_0$ predictions for each data point in the testing set were obtained and plotted in a histogram.

**1. Predicting wheter it rains:**

```{r predict rain}
#| echo: false
#| fig-cap: "Figure 2 - Histogram for the probability of not raining."
binary_rain_data <- test_df
binary_rain_data$y[binary_rain_data$y != 0] <- 1

#Estimate pi0 (probability of not raining)
raw_pred <- predict(model, newdata = binary_rain_data)
pi0_hat <- plogis(raw_pred[, 1])
hist(pi0_hat, main = "Probability of not raining", xlab = "Probability of not raining", ylab = "Frequency", col = "lightblue", border = "black")
```

The estimates of $\pi_0$ are concentrated around very high values, which agrees with the observation that in most days in the testing set it did not rain.

To assess the model numerically, we used the Brier score, defined by the formula

$$BS = \frac{1}{n}\sum_{i=1}^n(f_i-o_i)^2$$ where $f_i$ is the predicted probability, $o_i$ is the actual outcome (0 or 1) and n is the dataset size. In this case, n is the testing set size, which is 1200.


```{r}
# Brier Score
# The probability of raining (1 - pi_0) is expected to be pretty low, as most data points are 0
Brier_Score = mean(((1-pi0_hat) - binary_rain_data$y)^2)
print(Brier_Score)
```



We got a score equal to 0.139, a relatively small value, which indicates that the model performs reasonably well.

Conversely, to evaluate how well the model predicted the amount of rain in the days that rainfall was observed, we first subset the testing set to only rainy days.

We proceeded to calculate the mean rainfall from the $\alpha$ and $\beta$ parameters of the $Gamma$ distribution, per the following formula:

$$\mu = \frac{\alpha}{\beta}$$ These mean values were then plotted against the actual observed rainfall values:

```{r}
#| fig-cap: "Figure 3 - Estimated rainfall means vs true means."
#Subset for rainy days (y different from 0)
rain_test <- test_df[binary_rain_data$y != 0,]

rain_pred <- predict(model, newdata = rain_test)

alpha <- exp(rain_pred[,2])
beta <- exp(rain_pred[,3])

# Precipitation estimates
est_means <- alpha/beta

# True precipitation means
true_means <- rain_test$y

plot(true_means,est_means,pch=20, 
     xlab = "True means",
     ylab = "Estimated means")
```

The model seems to perform generally well considering that the larger values in the x-axis are few and sparse, with the larger cloud of points indicating a general agreement between predicted and real values, with the larger values exhibiting outlying behavior.

The root mean-squared error (RMSE) was used to diagnose the model's ability to predict the amount of rainfall. The RMSE is calculated via the following formula:

$$RMSE = \sqrt{\frac{1}{n}\sum(y_i-\hat{y_i})^2}$$ where $y_i$ are the observed values and $\hat{y_i}$ are the predicted values for rainfall, in this case.



```{r, include=FALSE}
RMSE = sqrt(mean((est_means - true_means)^2))
print(RMSE)
```
The resulting value for the RMSE was 3.93. Given the scale of the data, this value is not particularly low, however some reasonably outlying behavior was observed for some data points, which likely inflated some of the residuals, driving up the RMSE.

```{r, include=FALSE}
# n is the number of rainy days in the test set
  n <- sum(binary_rain_data$y != 0)
  alpha <- as.numeric(alpha)
  beta <- as.numeric(beta)
  i1 <- qgamma(0.05, alpha, beta)
  i2 <- qgamma(0.95, alpha, beta)
  coverage <- 0
```

```{r}
  for (j in 1:n){
    if(i1[j] < rain_test$y[j] & rain_test$y[j] < i2[j])
    coverage = coverage + 1}
    print(coverage/n)
```

The resulting value for the RMSE was 2.248. Given the scale of the data, this value is not particularly low, however some reasonably outlying behavior was observed for some data points, which likely inflated some of the residuals, driving up the RMSE.

Further into the rainfall predictions assessment, we built 90% confidence intervals for each of the data points and calculated their coverage, i.e, how many contain the true rainfall values. The resulting value was 0.922, indicating very high coverage, meaning that over 92% of the confidence intervals contained the true precipitation values.
:::

:::{.callout-note collapse="true"}

## Part 4: Scientific Questions

**Probability of no rainfall vs temperature**

To visually assess how temperature influences $\pi_0$, we plotted them against each other:

```{r}
#| echo: false
#| fig-cap: "Figure 4 - Temperature vs probability of no rainfall."
wholedata <- weather[year(weather$time) > 1960 & year(weather$time) < 2001,]
```


```{r Convert all temperature values from Kelvin (K) to degrees Celsius (°C)}
wholedata$tmax <- wholedata$tmax - 273.15
str(wholedata)
```

```{r}
wholey <- wholedata$pr
wholex <- wholedata$tmax
whole_df <- data.frame(x = wholex, y = wholey)
whole_pred <- predict(model, newdata = whole_df)
pi_whole <- plogis(whole_pred[, 1])
plot(pi_whole, whole_df$x, xlab = "Probability of no rainfall", ylab = "Temperature")
```

The plot indicates an increase in the probability of no rainfall as the temperature increases, which seems counter-intuitive...

**Rainfall quantiles vs temperature**

Now, plotting the temperature against the 5%, 50% and 95% quantiles of the rainfall distribution:

```{r}
#| echo: false
#| fig-cap: "Figure 5 - Temperature vs rainfall quantiles."
whole_alpha <- exp(whole_pred[, 2])
whole_beta <- exp(whole_pred[, 3])

# z is 1 for rainy days, 0 otherwise
z <- ifelse(wholedata$pr != 0, 1, 0)
rainyday <- wholedata$pr[z == 1]

int1 <- qgamma(0.05, whole_alpha[z == 1], whole_beta[z == 1])
int2 <- qgamma(0.50, whole_alpha[z == 1], whole_beta[z == 1])
int3 <- qgamma(0.95, whole_alpha[z == 1], whole_beta[z == 1])
plot(wholedata$tmax[z == 1], int1, main = "5%", ylab = "Rainfall", xlab = "Temperature")
plot(wholedata$tmax[z == 1], int2, main = "50%", ylab = "Rainfall", xlab = "Temperature")
plot(wholedata$tmax[z == 1], int3, main = "95%", ylab = "Rainfall", xlab = "Temperature")
```

The general trend in all three plots indicates that the amount of precipitation decreases when the temperature increase, with steeper increase for smaller amounts of rainfall. Only for high rainfall amounts (95% quantile) does the trend indicate an initial positive dependence between the variables, only to become negative afterwards as per the other two plots.

:::

:::{.callout-note collapse="true"}

## Appendix

Coming Soon!


Training Plots for the four models fitted to the training data. The blue line represents the training loss, while the red line represents the validation loss. The x-axis represents the number of epochs, while the y-axis represents the loss value.

```{r, eval=FALSE}
# call the png files saved in Part 2 to display the training plots for the four models fitted to the training data. The blue line represents the training loss, while the red line represents the validation loss. The x-axis represents the number of epochs, while the y-axis represents the loss value.
par(mfrow = c(2, 2))
model1_plot <- png::readPNG("model1_training_plot.png")
grid::grid.raster(model1_plot)
model2_plot <- png::readPNG("model2_training_plot.png")
grid::grid.raster(model2_plot)
model3_plot <- png::readPNG("model3_training_plot.png")
grid::grid.raster(model3_plot)
model4_plot <- png::readPNG("model4_training_plot.png")
grid::grid.raster(model4_plot)
```

:::
